But I can't preregister my research: Improving the reproducibility and transparency of ecology and conservation with adaptive preregistration for model-based research

### 2024-05-05 _analysis/ms.qmd
Add docx output format to ms
update discussion section of ms after discussion with libby

@@

![Insert caption. Number each step so i can refer to it in the headings and text below. Make it clear in the figure that we're not doing the build interface and conduct usability testing component of implement, nor the evaluate and iterate. we throw that out to the community of practice to do that part.](images/Pasted image 20240314123346.png){#fig-UCD}

@@

-    \[ \] #todo consider as #discussion point in this manuscript #writing

@@

## Preregistration Template

Blah what to say?

## Adaptive Preregistration Methodology

User research revealed critical features of the model development process, namely the inherent iterative nature of model development, and requirements for model checking that forced us to reconsider the procedure for completing a preregistration suitable to ecological modelling. Consequently, we took an ‘expanded view of preregistration’, “Adaptive Preregistration”. Key features: - plans for preregistering and deploying flexible analysis strategies: - Iterative, cyclical preregistration rather than linear process that maps onto the iterative cycles of model development generating - interim preregistrations marking key points in the analysis

@@

@@

**Resolution of the template**

@@

|-----------|-----------|----------------------------|-----------|-----------|

>>>>

  docx: default

@@

![Insert caption. Number each step so i can refer to it in the headings and text below. Make it clear in the figure that we're not doing the build interface and conduct usability testing component of implement, nor the evaluate and iterate. we throw that out to the community of practice to do that part.](images/Pasted%20image%2020240314123346.png){#fig-UCD}

@@

-   [ ] #todo consider as #discussion point in this manuscript #writing

@@

## Benefits, challenges, Limitations of Adaptive Preregistration

User research revealed inherent features of the model development process, namely its iterative workflow and adaptive decision making, were critical barriers to modellers adopting preregistration in their modelling practice. In the case of the VEFMAP modelling study, there were critical uncertainties about the the appropriate model specification given limitations of the underlying dataset which prevented the researchers from being able to pre-register fully specified models without undertaking some exploratory analysis. Such practices would preclude the application of preregistration to this study because these decisions would violate preregistration's core principle of *data-independent* decision-making.

*What would have happened without preregistration / adaptive preregistration??* (insert into last sentence of above para)

-   In the absence of any preregistration, adaptive or not, (#todo check this with Jian, because maybe his practices are different), modellers would typically explore the entire dataset
-   Explain what is wrong with this, what the risk of this is.

*Next para:*

To mitigate the risk of researcher degrees of freedom and to try to preserve some degree of data-independence in decision-making we combined data-partitioning into the adaptive preregistration process and split the preregistration process into two phases, resulting in three interim preregistrations. We first preregistered a preliminary or 'pilot' analysis, in which the researchers under took directed exploration of a subset of the dataset. The pilot analysis helped the researchers to understand what models could feasibly be fitted for their two outcome variables given uneven spatio-temporal density and patchiness of the dataset. Although the pilot analysis was 'exploratory' in nature, by preregistering the preliminary analysis, any exploration occurring here was still *directed* in that the scope of the pilot analysis was bounded and determined by transparently explicated analysis goals and articulartion of critical uncertainties. Following preregistration of the pilot analysis, the researchers then switched back to the *analysis* phase of *adaptive preregistration*. Carrying out the preregistered pilot analysis. The results of the pilot analysis aided in determining the *main analysis* which would be conducted on the full dataset and preregistered using the EcoEvoModelling template.

Researchers preregistered flexibility into their main analysis, because despite the pilot analysis, some critical uncertainties still reamined cf. limitations on the data.

-   outline the uncertainties: model family, and something else. see results.

Explain the different models, and the different

-   loosely preregistered (Baldwin) grounds for interpretation made explicit

Consequently, we took an ‘expanded view of preregistration’, “Adaptive Preregistration”. Key features:

\- plans for preregistering and deploying flexible analysis strategies:

\- Iterative, cyclical preregistration rather than linear process that maps onto the iterative cycles of model development generating

\- interim preregistrations marking key points in the analysis

@@

**Resolution of the template - tradeoff between meeting preregistration objectives, and burden and benefits to the researcher.**

*This will lead in nicely to the final criticism*

@@

@@

|-------------|-------------|----------------------|-------------|-------------|
